The relative roles of genetic, ecological and environmental factors in species radiations

### 2018-01-11 paper/draft.tex
additions to intro

One of the first and to this day highly influential formal theories of island biodiversity is MacArthur and Wilson's seminal equilibrium theory of island biogeography (ETIB). %TODO: ref
In it, species richness on islands is the result of opposing rates of immigration and extinction, both of which in turn are modified by island size (area) and its distance from the mainland (isolation).
This relationship was since successfully applied to a multitude of islands %TODO: refs
, but also island like systems, such as coral reefs, lakes/ponds or forest fragments. %TODO: refs
This general applicability, however, cannot hide the fact that the ETIB fails to deliver a mechanistic explanation of the processes at play.
It has particularly been criticized for its lack of considering speciation. %TODO: refs

@@

The problem of investigating these processes is the time frame needed for a proper evaluation.
Because islands biogeography takes into account evolutionary time frames (millions of years),
comprehensive and meaningful data is difficult to obtain.
In a recent review \cite{leidinger2017biodiversity}, we analyzed a powerful alternative to field experiments in island biogeography
--- mechanistic (/simulation) models.
Unfortunately, most island models failed to include more explicit genetic or micro evolution.
Instead, some models featured a concept known as ``protracted speciaton'' \cite{rosindellXXX}, where speciation occurs with a delay from the actual
divergence event if the population remains isolated during that time.
But nature is usually more complex than this and genetic mechanisms have important ramifications for speciation/radiation and species richness.
[EXAMPLES!]

@@

In the following, we present a model that remedies both of the adressed shortcomings by implementing genetic processes on a niche-explicit background
in an island biogeography framework.
We will adress the questions of what relative impact does (a) genetic linkage and (b) reproductive isolation have on emerging species numbers and diversity,
and how (c) landscape heterogeneity (habitat niche diversity) and (d) topographical complexity differently affect radiatons.
[further questions thus tackled: traits of colonizer species, ...]

>>>>

Recently, a group of influential island biogeography researchers proposed a set of 50 questions in island biogeography that still need answering (Patiño et al. 2017).
A subset of these questions deals with 
One of the first and to this day highly influential formal theories of island biodiversity is MacArthur and Wilson's seminal equilibrium theory of island biogeography (ETIB). %TODO: ref
In it, species richness on islands is the result of opposing rates of immigration and extinction, both of which in turn are modified by island size (area) and its distance from the mainland (isolation).
This relationship was since successfully applied to a multitude of islands %TODO: refs
, but also island like systems, such as coral reefs, lakes/ponds or forest fragments. %TODO: refs
This general applicability, however, cannot hide the fact that the ETIB fails to deliver a mechanistic explanation of the processes at play.
It has particularly been criticized for its lack of considering speciation. %TODO: refs

@@

The problem of investigating these processes is the time frame needed for a proper evaluation.
Because islands biogeography takes into account evolutionary time frames (millions of years),
comprehensive and meaningful data is difficult to obtain.
In a recent review \cite{leidinger2017biodiversity}, we analyzed a powerful alternative to field experiments in island biogeography
--- mechanistic (/simulation) models.
Unfortunately, most island models failed to include more explicit genetic or micro evolution processes.
% reasons: underlying theories: Recently, a group of influential island biogeography researchers proposed a set of 50 questions in island biogeography that still need answering (Patiño et al. 2017).
%A subset of these questions deals with 
Instead, some models featured a concept known as ``protracted speciaton'' \cite{rosindellXXX}, where speciation occurs with a delay from the actual
divergence event if the population remains isolated during that time.
But nature is usually more complex than this and genetic mechanisms have important ramifications for speciation/radiation and species richness.
[EXAMPLES!]

@@

In the following, we present a model that remedies both of the adressed shortcomings by implementing genetic processes on a niche-explicit background
using an island biogeography framework.
We will adress the questions of what relative impact does (a) genetic linkage and (b) reproductive isolation have on emerging species numbers and diversity,
and how (c) landscape heterogeneity (habitat niche diversity) and (d) topographical complexity differently affect radiatons.
[further questions thus tackled: traits of colonizer species, ...]
